### 2023-09-19 doc/_logistic_sim.qmd
implement grammar suggestiosn

The inference method is based on the model that assumes that the time passed
between dilutions $\tau \approx 8$ generations, the change in frequency for a
mutant barcode can be approximated from cycle $t$ to the next cycle $t + 1$ as
$$
f_{t+1}^{(m)} = f_t^{(m)} \mathrm{e}^{(s^{(m)} - \bar{s}_t)\tau},
$${#eq-logistic_fitness_fn}
where $s^{(m)}$ is the relative fitness for strain $i$ compared to the ancestral
strain and $\bar{s}_t$ is the mean fitness of the population at cycle $t$. We
implemented a numerical experiment following the logistic growth model described
in @eq-logistic_fn to test this assumption. @fig-SIX_logistic_growth_01 shows an
example of the deterministic trajectories for 50 labeled neutral lineages and
1000 lineages of interest. The upper red curve that dominates the culture 
represents the unlabeled ancestral strain included in the experimental design
described in @sec-experiment.

@@

To simulate multiple growth-dilution cycles, we take the population composition
at the final time point and use it to initialize a new logistic growth
simulation. @fig-SIX_logistic_growth_02 shows the resulting number of cells at
the final time point of a cycle over multiple growth-dilution cycles for the
genotypes in @@fig-SIX_logistic_growth_01. We can see that the adaptive lineages
(blue curves) increase in abundance, while detrimental lineages (red curves)
decrease.

@@

In @sec-fitness_model we derive the functional form to infer the relative 
fitness of each lineage as
$$
\frac{1}{\tau}\ln \frac{f_{t+1}^{(b)}}{f_{t}^{(b)}} = (s^{(b)} - \bar{s}_t).
$${#eq-logistic_logfreq}
@fig-SIX_logistic_growth_03 shows the corresponding log frequency ratio curves
for the logistic growth simulation. The displacement of these curves with 
respect to the neutral lineages determines the ground truth relative fitness
value for these simulations.

@@

1. Poisson noise between dilutions. For this, we take the final point of the
logistic growth simulation and sample a random Poisson number based on this
final point to set the initial condition for the next cycle.
2. Gaussian noise when performing the measurements. When translating the 
underlying population composition to the number of reads, we can add a custom 
amount of Gaussian noise.

>>>>

The inference method is based on the model that assumes that the time passed
between dilutions $\tau \approx 8$ generations, the change in frequency for a
mutant barcode can be approximated from cycle $t$ to the next cycle $t + 1$ as
$$
f_{t+1}^{(m)} = f_t^{(m)} \mathrm{e}^{(s^{(m)} - \bar{s}_t)\tau},
$${#eq-logistic_fitness_fn}
where $s^{(m)}$ is the relative fitness for strain $i$ compared to the ancestral
strain and $\bar{s}_t$ is the mean fitness of the population at cycle $t$. To
test this assumption, we implemented a numerical experiment following the
logistic growth model described in @eq-logistic_fn. @fig-SIX_logistic_growth_01
shows an example of the deterministic trajectories for 50 labeled neutral
lineages and 1000 lineages of interest. The upper red curve that dominates the
culture represents the unlabeled ancestral strain included in the experimental
design described in @sec-experiment.

@@

To simulate multiple growth-dilution cycles, we take the population composition
at the final time point and use it to initialize a new logistic growth
simulation. @fig-SIX_logistic_growth_02 shows the resulting number of cells at
the last time point of a cycle over multiple growth-dilution cycles for the
genotypes in @@fig-SIX_logistic_growth_01. We can see that the adaptive lineages
(blue curves) increase in abundance, while detrimental lineages (red curves)
decrease.

@@

In @sec-fitness_model, we derive the functional form to infer the relative 
fitness of each lineage as
$$
\frac{1}{\tau}\ln \frac{f_{t+1}^{(b)}}{f_{t}^{(b)}} = (s^{(b)} - \bar{s}_t).
$${#eq-logistic_logfreq}
@fig-SIX_logistic_growth_03 shows the corresponding log frequency ratio curves
for the logistic growth simulation. The displacement of these curves with 
respect to the neutral lineages determines the ground truth relative fitness
value for these simulations.

@@

1. Poisson noise between dilutions. For this, we take the final point of the
logistic growth simulation and sample a random Poisson number based on this
last point to set the initial condition for the next cycle.
2. Gaussian noise when performing the measurements. When translating the 
underlying population composition to the number of reads, we can add a custom 
amount of Gaussian noise.
